---
title: "01_12_2017"
author: "Ruijuan Li"
date: "1/12/2017"
output: html_document
---

# read the manual 
# Bring remaining questions about BRMS.  JM will try to answer.  If you wished that you had read the manual or help files before last week, do so before this week...

```{r}
library(brms)
vignette('brms_overview')
vignette("brms_overview")

# 1) 
# something about setup 
# rstan_options (auto_write=TRUE)
# options (mc.cores=parallel::detectCores ()) # Run on multiple cores
# a general setup thing? 

# 2) 
# explain the result? 
# group-level effect; population level effect; family specific parameters 
# sd_shelf_intercept is not normally distributed... what is this about? 

# 3) about shiny 
# the usage and application of shiny in bioinformatics 

# 4) about fixed & random effect 
# when to determine an effect should be fixed or random 
# when to use unique intercept? # use tomato dataset as and example 

# 5) hypothesis testing 
# don't know how to interpret the result, which is significant, which is not? 

# 6) as.factor(tank) in homework Q2 
# maybe need all predictors to be factors? 

# 7) in Q3, your map2stan model didn't include trt effect. 
```

# homework, Fit and interpret a BRMS model to your own data set of your choice.  Be prepared to explain you analysis.
# fit my rapa data using brms 
```{r}
GC.pheno <- read.csv("/Users/ruijuanli/Desktop/Brassica_project/RNA_seq/data/shade_avoidance_expt_in_GC.csv",na.string=c("NA","N/A")) # replace all NA w/ N/A (the standard missing data format in R)
GH.pheno <- read.csv("/Users/ruijuanli/Desktop/Brassica_project/RNA_seq/data/shade_avoidance_expt_in_GH.csv", na.string=c("NA","N/A"))

head(GC.pheno)
head(GH.pheno)

summary(GC.pheno) 
summary(GH.pheno)
```

# reformat data, get a factor for rep 
```{r}
library("stringr") # to help substring letters use negative to count backwards

# GC 
names(GC.pheno)[3] <- "gt"
GC.pheno$rep <- as.factor(str_sub(GC.pheno$barcode,-2,-2))  # 8 reps 

# 
GC.pheno$height <- GC.pheno$No..of.internodes * GC.pheno$intleng
head(GC.pheno)

GC.pheno.data <- GC.pheno[, c("gt", "trt", "leaves", "No..of.internodes", "leafl", "leafw", "petleng", "rep", "height")]
colnames(GC.pheno.data) <- c("gt", "trt", "leaf_No", "internode_No", "leafl_ave", "leafw_ave", "petleng_ave", "rep", "height")
head(GC.pheno.data)

# GH 
GH.pheno$rep <- as.factor(GH.pheno$rep)
GH.pheno$rep 
head(GH.pheno)

GH.pheno$rep <- as.factor(GH.pheno$rep)
GH.pheno$rep 
head(GH.pheno)

GH.pheno$petleng_ave <- round(rowMeans(subset(GH.pheno, select=c("pet1", "pet2", "pet3")), na.rm = T), digits = 1)
GH.pheno.data <- GH.pheno[, c("trt", "rep", "intleng", "leafleng", "leafwid", "leafnum", "internum", "species", "petleng_ave")]
colnames(GH.pheno.data) <- c("trt", "rep", "height", "leafl_ave", "leafw_ave", "leaf_No", "internode_No", "gt", "petleng_ave")
head(GH.pheno.data)
GH.pheno.data$trt
``` 

# fit the data using BRMS model 
```{r}
# build models for plant height, leaf length average, and leaf width average 
GC.pheno.data$trt <- relevel(GC.pheno.data$trt, ref = "SUN")
GC.pheno.data$gt <- relevel(GC.pheno.data$gt, ref = "R500") 

m.GC_height <- brm(height ~ trt * gt + (1|rep), 
            data = GC.pheno.data,
            prior = c(set_prior("normal(0,10)",class="b")) # sets prior for all b coefficients  
) 

summary(m.GC_height)
fixef(m.GC_height)

m.GC_leafl <- brm(leafl_ave ~ trt * gt + (1|rep), 
            data = GC.pheno.data,
            prior = c(set_prior("normal(0,10)",class="b")) # sets prior for all b coefficients  
) 

summary(m.GC_leafl)
fixef(m.GC_leafl) 

m.GC_leafw <- brm(leafw_ave ~ trt * gt + (1|rep), 
            data = GC.pheno.data,
            prior = c(set_prior("normal(0,10)",class="b")) # sets prior for all b coefficients  
) 

summary(m.GC_leafw)
fixef(m.GC_leafw)  

### GH 
GH.pheno.data$trt <- relevel(GH.pheno.data$trt, ref = "ND")
GH.pheno.data$gt <- relevel(GH.pheno.data$gt, ref = "R500") 

m.GH_height <- brm(height ~ trt * gt + (1|rep), 
            data = GH.pheno.data,
            prior = c(set_prior("normal(0,10)",class="b")) # sets prior for all b coefficients  
) 

summary(m.GH_height)
fixef(m.GH_height)

m.GH_leafl <- brm(leafl_ave ~ trt * gt + (1|rep), 
            data = GH.pheno.data,
            prior = c(set_prior("normal(0,10)",class="b")) # sets prior for all b coefficients  
) 

summary(m.GH_leafl)
fixef(m.GH_leafl) 

m.GH_leafw <- brm(leafw_ave ~ trt * gt + (1|rep), 
            data = GH.pheno.data,
            prior = c(set_prior("normal(0,10)",class="b")) # sets prior for all b coefficients  
) 

summary(m.GH_leafw)
fixef(m.GH_leafw)  
```

# test the effect of gt, trt, and gt:trt for each trait 
```{r}
### GC 
summary(m.GC_height)
# height 
hypothesis(m.GC_height, "gtIMB211=0")
hypothesis(m.GC_height, "trtSHADE=0")
hypothesis(m.GC_height, "trtSHADE:gtIMB211=0")
# leafl
hypothesis(m.GC_leafl, "gtIMB211=0")
hypothesis(m.GC_leafl, "trtSHADE=0")
hypothesis(m.GC_leafl, "trtSHADE:gtIMB211=0")

hypothesis(m.GC_leafw, "gtIMB211=0")
hypothesis(m.GC_leafw, "trtSHADE=0")
hypothesis(m.GC_leafw, "trtSHADE:gtIMB211=0")

### GH 
summary(m.GH_height)
# height 
hypothesis(m.GH_height, "gtIMB211=0")
hypothesis(m.GH_height, "trtD=0")
hypothesis(m.GH_height, "trtD:gtIMB211=0")
# leafl 
hypothesis(m.GH_leafl, "gtIMB211=0") 
hypothesis(m.GH_leafl, "trtD=0")
hypothesis(m.GH_leafl, "trtD:gtIMB211=0")
# leafw 
hypothesis(m.GH_height, "gtIMB211=0")
hypothesis(m.GH_height, "trtD=0")
hypothesis(m.GH_height, "trtD:gtIMB211=0") 
```

# output as figure 
```{r}
library(ggplot2)
### GC 
summary(m.GC_height)    
fixef(m.GC_height)
fixef(m.GC_leafl)
fixef(m.GC_leafw) 

### GH 
fixef(m.GH_height)
fixef(m.GH_leafl)
fixef(m.GH_leafw)
``` 



