---
output: pdf_document
---
# Statistical Rethinking Chapter 4 problems

__Name: Ruijuan Li


# For 03/17/16

## 4E1 

the first one is the likelihood

## 4E2 

two, mu and sigma

## 4E3 

this formula should look very similar to the one on P83 

## 4M1 for the model definition below, simulate observed heights from the prior. 
```{r}
library(rethinking)
sample_mu <- rnorm(1e4, 0, 10)
sample_sigma <- runif(1e4, 0, 10)
prior <- rnorm(1e4, sample_mu, sample_sigma)
dens(prior)
```

## 4M2 translate the model into a map formula
```{r}
data("Howell1")
d <- Howell1
d <- d[d$age >=18,]
m <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu ~ dnorm(0, 10),
    sigma ~ dunif(0, 10)
  ), 
  data = d # prblem... 
)
```

# For 03/24/16

## 4E4

second line

## 4E5

three, alpha, beta, and sigma

## 4M3

yi ~ Normal(mui, sigma)

mui = a+b*xi 

a ~ Normal(0, 50)

b ~ Normal(0, 10)

sigma ~ Uniform(0, 50)

## 4M4

hi ~ Normal(mui, sigma) 

mui = a + b*xi 

a ~ Normal(138, 50)

b ~ Uniform(6, 12)

sigma ~ Uniform(0, 50)

## 4M5
138 --> 120

## 4M6

variance of 64 means a sd of 8, so 50 --> 8

# For 03/31/16

## 4H1 predicted height or mean of height??? 

```{r}
# import data
library(rethinking)
data("Howell1")
d <- Howell1

# model fit, normal linear regression 

model1 <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b*weight, 
    a ~ dnorm(178, 100),
    b ~ dnorm(0, 10),
    sigma ~ dunif(0, 50)
  ), 
  data=d)

# plot the model
plot(height ~ weight, data=d)
abline(a=coef(model1)["a"], b = coef(model1)["b"])
# extract samples to form a posterior
post <- extract.samples(model1)
colnames(post)

# predict average height values
mu.link <- function(weight) post$a + post$b*weight
weight.seq <- c(46.95, 43.72, 64.78, 32.59, 54.63)
mu <- sapply(weight.seq, mu.link)
mu.mean <- apply(mu, 2, mean)
mu.mean
mu.HPDI <- apply(mu, 2, HPDI, prob=0.89) 
mu.HPDI

# predicted height:  158.2752 152.5763 189.7340 132.9389 171.8256
# 89% HPDI interval 
# [,1]     [,2]     [,3]     [,4]     [,5]
# |0.89 157.4944 151.8958 188.3771 132.2422 170.7728
# 0.89| 159.1143 153.3666 191.2324 133.5591 172.8721
```

## 4H2
```{r}
# get data
d2 <- d[d$age < 18, ]
nrow(d2)
```

# (a)
```{r}
# fit a linear regression model
model2 <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b*weight,
    a ~ dnorm(138, 100),
    b ~ dnorm(0, 10),
    sigma ~ dunif(0, 50)
  ), 
data=d2)

# plot
plot(height ~ weight, data=d2)
abline(a=coef(model2)["a"], b = coef(model2)["b"])

# extract slope values b
post2 <- extract.samples(model2)
dens(post2$b*10)
HPDI(post2$b*10)
mean(post2$b)*10
# [1] 27.18586 
# |0.89    0.89| 
# 26.05893 28.25237 
```

# (b)
```{r}
plot(height ~ weight, data=d2)

# use link to generate distribution of posterior values for mu
mu.link <- function(weight) post2$a + post2$b*weight
weight.seq <- seq(0, 40, by=1)
mu <- sapply(weight.seq, mu.link)
mu.mean <- apply(mu, 2, mean)
mu.HPDI <- apply(mu, 2, HPDI, prob=0.89)

# plot the distribution, the predicted mean, and the 89% interval 
plot(height ~ weight, data=d2, col=col.alpha(rangi2, 0.5))
lines(weight.seq, mu.mean)
shade(mu.HPDI, weight.seq)

# 89% interval for the predicted heights
sim.height <- sim(model2, data = list(weight=weight.seq))
height.PI <- apply(sim.height, 2, PI, prob=0.89)
shade(height.PI, weight.seq)
```

# (c)
```{}
the model doesn't fit the current data very well, because it seems children with weight below 10kg

and above 35kg had a slower growth rate as compared to children with weight between 10 to 35. While 
I made the assumption of steady growth rate, a polymonial regression with a squred curve may fit

the current data better and can possibly serve as a better model. 
```

## 4H3

# (a)
```{r}
# import data
d <- Howell1
nrow(d)

# fit model 
d$weight.log <- log(d$weight) # cannot standadize, produce 0, prohibit log transformation
model3 <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu <- a + b * weight.log,
    a~ dnorm(178, 100),
    b ~ dnorm(0, 100),
    sigma ~ dunif(0, 50)
  ),
data = d)

plot(d$height, d$weight.log)
```

# (b)
```{r}
plot(height ~ weight, data=Howell1,
col=col.alpha(rangi2, 0.4))
```

```{}
# one way for the rest #######################################
# use link to generate distribution of posterior values for mu
post3 <- extract.samples(model3)
mu.link <- function(weight) post3$a + post3$b*log(weight)
weight.seq <- seq(0, 70, by=1)
mu <- sapply(weight.seq, mu.link)
mu.mean <- apply(mu, 2, mean)
mu.HPDI <- apply(mu, 2, HPDI, prob=0.89)

# plot the distribution, the predicted mean, and the 89% interval 
lines(weight.seq, mu.mean)
shade(mu.HPDI, weight.seq)

# 89% interval for the predicted heights
sim.height <- sim(model3, data = list(weight=weight.seq))
height.PI <- apply(sim.height, 2, PI, prob=0.89)
shade(height.PI, weight.seq)

### a second way for the rest ####################################
weight.seq <-seq(0, 70, by=1)
pred_dat <- list(weight = weight.seq, weight.log=log(weight.seq))
mu <- link(model3, data = pred_dat) 

mu.mean <- apply(mu, 2, mean)
mu.PI <- apply(mu, 2, PI, prob=0.89)
sim.height <- sim(model3, data = pred_dat)
height.PI <- apply(sim.height, 2, PI, prob=0.89)

plot(height ~ weight.s, d, col=col.alpha(rangi2, 0.5))
lines(weight.seq, mu.mean)
shade(mu.PI, weight.seq)
shade(height.PI, weight.seq)
```
