---
output: html_document
---
# Statistical Rethinking Chapter 4 problems

__Name: Ruijuan Li


# For 03/17/16

## 4E1 

the first one is the likelihood

## 4E2 

two, mu and sigma

## 4E3 

this formula should look very similar to the one on P83 

## 4M1 for the model definition below, simulate observed heights from the prior. 
```{r}
library(rethinking)
sample_mu <- rnorm(1e4, 0, 10)
sample_sigma <- runif(1e4, 0, 10)
prior <- rnorm(1e4, sample_mu, sample_sigma)
dens(prior)
```

## 4M2 translate the model into a map formula
```{r}
data("Howell1")
d <- Howell1
d <- d[d$age >=18,]
m <- map(
  alist(
    height ~ dnorm(mu, sigma),
    mu ~ dnorm(0, 10),
    sigma ~ dunif(0, 10)
  ), 
  data = d # prblem... 
)
```

# For 03/24/16

## 4E4

second line

## 4E5

three, alpha, beta, and sigma

## 4M3

yi ~ Normal(mui, sigma)

mui = a+b*xi 

a ~ Normal(0, 50)

b ~ Normal(0, 10)

sigma ~ Uniform(0, 50)

## 4M4

hi ~ Normal(mui, sigma1) 

mui = a + b*xi 

a ~ Normal(mu2, sigma2)

b ~ Normal(mu3, sigma3)

sigma ~ Uniform(n, m)

## 4M5

mu2 = 120

b > 0

## 4M6

sigma <= 64

sigma2 <= 64

# For 03/31/16

## 4H1

## 4H2

## 4H3
